---
title: "PA1_template"
author: "Wilma"
date: "2 december 2017"
output: html_document:
  keep_md: true
---

* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Loading and preprocessing the data

```{r}
setwd("~/GitHub/RepData_PeerAssessment1")
activity <- read.csv("activity.csv")
summary(activity)
str(activity)

# Turn the date data into a valid date class to allows for easier processing : Dates are in YYYY-MM-DD format
dates <- strptime(activity$date, "%Y-%m-%d")
activity$date <- dates

# Keep a list of all possible days
uniqueDates <- unique(dates)

# Keep a list of all possible intervals
uniqueIntervals <- unique(activity$interval)

```


## 2. What is mean total number of steps taken per day?
### 2.1 Total number of steps per day

```{r}
stepsSplit <- split(activity$steps, dates$yday)
sum_by_date <- sapply(stepsSplit, sum, na.rm=TRUE)
sum_by_date

```
### 2.2 Histogram

```{r}
plot(uniqueDates, sum_by_date, main="Histogram of Total Steps per day", xlab="Days in October to November 2012", ylab="Frequency", type="h", lwd=4, col=blues9)


```


### 2.3 Mean of total number of steps taken per day

```{r}
mean_by_date <- sapply(stepsSplit, mean, na.rm=TRUE)
mean_df <- data.frame(date=uniqueDates, mean_by_date=mean_by_date, row.names=NULL)
mean_df

```

### 2.4 Median of total number of steps per day
```{r}
median_by_date <- sapply(stepsSplit, median, na.rm=TRUE)
median_df <- data.frame(date=uniqueDates, median_by_date=median_by_date, row.names=NULL)
median_df


```


## 3. What is the average daily activity pattern?

### 3.1 Make a time series plot

```{r}
intervals <- split(activity$steps, activity$interval)
avg_Interval <- sapply(intervals, mean, na.rm=TRUE)

plot(uniqueIntervals, avg_Interval, type="l",main="Avg number of Steps per Interval - All Days", xlab="Interval",    ylab="Average No of Steps", lwd=2, col="green")


 


```

### 3.2 Interval with most steps

```{r}

maxIntervalDays <- max(avg_Interval, na.rm=TRUE)
maxIndex <- as.numeric(which(avg_Interval == maxIntervalDays))
maxInterval <- uniqueIntervals[maxIndex]
maxInterval

```




## 4. Imputing missing values

###4.1 Calculate and report the total number of missing values in the dataset

```{r}
nulls <- is.na(activity$steps)
sum(nulls)
```

### 4.2 Fill in missing values

```{r}
mean_by_date[is.nan(mean_by_date)] <- 0
summary(mean_by_date)
meancolumn <- rep(mean_by_date, 288)

Steps <- activity$steps
null_steps <- is.na(Steps)
summary (null_steps)

Steps[null_steps] <- meancolumn[null_steps]

```

### 4.3 Create a new dataset

```{r}
activity2 <- activity
summary(activity2)
activity2$steps <- Steps
summary(activity2)


```

### 4.4 Make a histogram of the total number of steps taken each day  
```{r}
stepsSplit2 <- split(activity2$steps, dates$yday)
 stepsSplit2 <- split(activity2$steps, dates$yday)
sum_by_date2 <- sapply(stepsSplit2, sum)
sum_by_date2

plot(uniqueDates, sum_by_date2, main="Histogram of Total Steps per day(After Imputing)", xlab="Days in October to November 2012", ylab="Frequency", type="h", lwd=4, col=blues9)

```


### 4.5 Calculate and report the mean and median total number of steps taken per day. 

#### 4.5.1 Mean recalculated

```{r}
mean_by_date2 <- sapply(stepsSplit2, mean)
mean_df2 <- data.frame(date=uniqueDates, mean_by_date = mean_by_date,mean_by_date2 = mean_by_date2 , row.names=NULL)
mean_df2

```

#### 4.5.2 Median recalculated
```{r}
median_by_date2 <- sapply(stepsSplit2, median)
median_df2 <- data.frame(date=uniqueDates, median_by_date = median_by_date, median_by_date2 = median_by_date2, row.names=NULL)
median_df2

```

#### 4.5.3 Total number of steps per day recalculated

```{r}
sum_by_date2_df <- data.frame(date = uniqueDates, sum_by_date= sum_by_date, sum_by_date2=sum_by_date2, row.names=NULL)
sum_by_date2_df
```


### 4.6 Do these values differ from the estimates from the first part of the assignment?

After replacing the days with all '0', their mean values changed. 

After replacing the days with all 'NA', their median values changed.

For the other days nothing changed.



### 4.7 What is the impact of imputing missing data on the estimates of the total daily number of steps?

After replacing the days with all '0', their total daily values changed. 

For the other days nothing changed.




## 5. Are there differences in activity patterns between weekdays and weekends?

### 5.1 Create a new factor variable in the dataset with two levels - "weekday" and "weekend" 

```{r}

weekdays <- dates$wday
daytype <- rep(0, length(weekdays)-1)
daytype[weekdays >= 1 & weekdays <= 5] <- 1
daytype[weekdays == 6 | weekdays == 0] <- 2
day_factor <- factor(daytype, levels=c(1,2), labels=c("Weekday", "Weekend"))
activity2$type_of_day <- day_factor
summary(activity2)
activity2_Weekday <- activity2[activity2$type_of_day == "Weekday", ]
activity2_Weekend <- activity2[activity2$type_of_day == "Weekend", ]


```
### 5.2 Make a panel plot containing a time series plot (i.e. type = "l") of the5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
activity2_weekday_split <- split(activity2_Weekday$steps, activity2_Weekday$interval)
activity2_weekend_split <- split(activity2_Weekend$steps, activity2_Weekend$interval)
meanSteps_Weekday <- sapply(activity2_weekday_split, mean)
meanSteps_Weekend <- sapply(activity2_weekend_split, mean)

 par(mfcol=c(2,1))
plot(uniqueIntervals, meanSteps_Weekday, type="l", main="Avg number of Steps per Interval - Weekdays", 
 xlab="Interval", ylab="Average No of steps", lwd=2, col="darkgreen")
plot(uniqueIntervals, meanSteps_Weekend, type="l", main="Avg number of Steps per Interval - Weekends", 
 xlab="Interval", ylab="Average No of steps",lwd=2, col="green")
 

```

## End of Report



